---
title: "Task3 -- Happiness vs Mood of Songs"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
df <- read.csv("happiness_w_mood_R/happiness_w_mood_R.csv")
df$developed_country = as.logical(df$developed_country)
df[is.na(df)] <- 0
```

# Data Exploration

Variables: - year - country - positive: number of positive songs -
neutral: number of neutral songs - negative: number of negative songs -
developed_country: indicator, true if the country is considered as
developed country

```{r}
summary(df)

par(mfrow = c(2, 2))
plot_happiness <- function(data, country_name) {
  country_data <- data[data$country == country_name, ]
  # output as individual plots
  #png(paste("plots/Happiness_in_", country_name, "_Over_Time.png", sep = ""),
  #    width = 600, height = 400)
  plot(
    country_data$year, 
    country_data$happiness, 
    type = "l", 
    main = paste("Happiness Score in", country_name, "Over Time"), 
    xlab = "Year", 
    ylab = "Happiness Score",
    col = "blue",
    lwd = 3,
    ylim = c(4.3, 7.5)
  )
  
  points(country_data$year, country_data$happiness, col = "blue", pch = 16)
  #dev.off()
}

for (c in unique(df$country)){
  plot_happiness(df, c)
}


# Rescale
# multiply by 10 to match the scale of the happiness score
df$positive <- df$positive / 110 * 10
df$neutral <- df$neutral / 110 * 10
df$negative <- df$negative / 110 * 10
```

```{r}
# Correlation matrix
cor_data <- df[, c("happiness", "positive", "neutral", "negative")]
cor_matrix <- cor(cor_data)
print(cor_matrix)

library(ggplot2)
library(corrplot)
#png("plots/complete_corr_heat.png", width = 600, height = 400)
corrplot(cor_matrix, method = "color")
#dev.off()
```

-   We see that `positive` and `negative` are strongly negatively
    correlated, which can cause multicollinearity issues in the
    regression models. Hence, we combine `postive` and `negative` into a
    new feature to avoid such issues.
-   Also, notice that happiness **is not correlated** with the count of
    positive/neutral/negative song in general.

```{r}
# higher value of comb means more positive songs were listened
df$comb <- df$positive - df$negative


# re-examine correlation 
cor_data <- df[, c("happiness", "neutral", "comb")]
cor_matrix <- cor(cor_data)
print(cor_matrix)
```

No highly correlated variables!

```{r}
par(mfrow = c(1, 2))

plot(df$comb, df$happiness, 
     xlab = "Combined Mood",
     ylab = "Happiness",
     col = "red")

plot(df$neutral, df$happiness, 
     xlab = "Neutral Mood",
     ylab = "Happiness",
     col = "red")
```

## Trend over Time by Country

```{r}
par(mfrow = c(1,2))

plot_country_trends <- function(data) {
  countries <- unique(data$country)

  for (country in countries) {
    country_data <- data[data$country == country, ]
    
    plot(
      country_data$year, 
      country_data$happiness, 
      type = "l", 
      main = paste("Trends for", country),
      xlab = "Year", 
      ylab = "Happiness Score", 
      col = "blue", 
      ylim = c(min(c(country_data$comb, country_data$neutral))
               , 8.5),
      lwd = 2
    )
    
    lines(country_data$year, country_data$neutral, col = "green", lwd = 2)
    lines(country_data$year, country_data$comb, col = "red", lwd = 2)
    
    legend(
      "topright", 
      legend = c("Happiness", "Neutral Moods", "Combined Moods"), 
      col = c("blue", "green", "red"), 
      lty = 1, 
      lwd = 2,
      cex = 0.5
    )
  }
}

plot_country_trends(df)
```

# Modelling

## Baseline Model

```{r}
m0 <- lm(happiness ~  comb + neutral, data = df)
summary(m0)
```

## Examine the effect of the development of a country

```{r}
boxplot(happiness~developed_country, main = "All Countries", col = c("pink", "lightblue"),
        data = df, xlab = "developed or not")
```

```{r}
fit0 <- lm(happiness ~ developed_country, data = df)
summary(fit0)
```

-   `developed_countryTRUE`: the difference in happiness between
    developed and developing countries on average is 0.867
-   we see from the boxplot that the IQR and mean of happiness score for
    developed countries are much higher than those of developing
    countries, which suggests the happiness score of developed countries
    are higher in general. Also, the `developed_countryTRUE` coefficient
    has a positive value of 0.867, which implies developed countries has
    higher happiness score. The null hypothesis $H_0:\beta_1 = 0$ has a
    p-value of 9.57e-11 \<\< 0.001, this means there is strong evidence
    against the null hypothesis, which aligns with our previous
    conclusion that developed countries has higher happiness score on
    average.

## Model with New Variable

```{r}
m1 <- lm(happiness ~ comb*developed_country +
           neutral*developed_country,
         data = df)
summary(m1)
```

```{r}
anova(m0, m1)
```

m1 is a better model, adding `developed_country` is reasonble.

## More Analysis on `devloped_country`

For better interpretability, we define a new variable `in_one` to
combine all 3 moods

```{r}
# higher value of in_one means people tend to listen to more positive songs
# lower value of in_one means people tend to listen to more negative songs
# 0.5 is a random weight
df$in_one <- df$positive - df$negative + 0.5 * df$neutral

# the final model
m3 <- lm(happiness ~ in_one*developed_country, data = df)
summary(m3)
plot(happiness ~ in_one, main = "Happiness Score vs Song Sentiment",
     xlab = "song sentiment", ylab = "happiness score",
     data = df[df$developed_country == TRUE,], col = "firebrick",
     ylim = c(4, 8), xlim = c(min(df$in_one), max(df$in_one)))
points(happiness ~ in_one, data = df[df$developed_country == FALSE,],
       col = "steelblue")
legend("topright", c("developed countries", "developing countries"), cex = 0.7,
       col = c("firebrick", "steelblue"), pch = 1)
coefs_hat <- coefficients(m3)
abline(coefs_hat[1], coefs_hat[2], lty = 1, lwd = 2)
abline(coefs_hat[1]+coefs_hat[3], coefs_hat[2]+coefs_hat[4], lty = 2, lwd = 2)
```

**Interpretation:**

-   We plot the linear fitted lines for developed and developing
    countries separately. Notice that there is a more obvious trend for
    developing countries while there is no obvious relationship between
    the sentiments of songs and the happiness score of one country for
    developed countries.
-   Notice that the R-squared value for the model is low, which suggests
    happiness score is not well-explained by the variables we used.

Additionally, we examine the correlation between happiness and sentiment
of songs separately for developed and developing countries.

```{r}
# General correlation 
cor_data <- df[, c("happiness", "in_one")]
cor_matrix <- cor(cor_data)
print(cor_matrix)
#png("plots/all_corr_heat.png", width = 600, height = 400)
corrplot(cor_matrix, method = "color")
#dev.off()
```

```{r}
# for developed countries
developed_df <- df[df$developed_country == TRUE, ]
cor_data <- developed_df[, c("happiness", "in_one")]
cor_matrix <- cor(cor_data)
print(cor_matrix)
#png("plots/developed_corr_heat.png", width = 600, height = 400)
corrplot(cor_matrix, method = "color")
#dev.off()
```

```{r}
# for developing countries
developing_df <- df[df$developed_country == FALSE, ]
cor_data <- developing_df[, c("happiness", "in_one")]
cor_matrix <- cor(cor_data)
print(cor_matrix)
#png("plots/developing_corr_heat.png", width = 600, height = 400)
corrplot(cor_matrix, method = "color")
#dev.off()
```

Based on the three correlation matrices, we can conclude that, overall,
there is no significant relationship between a country's happiness score
and the moods of songs its people prefer. A similar observation holds
true for developed countries. However, in developing countries, a
moderate negative correlation exists, indicating that individuals in
happier countries tend to listen to songs with more negative moods.

```{r}
# plotting for BI

par(mfrow = c(2,2))
plot_country_trends <- function(data) {
  countries <- unique(data$country)

  for (country in countries) {
    country_data <- data[data$country == country, ]
    # output the happiness and mood over time plots as individual files
    #png(paste("plots/Trends_for_", country, ".png", sep = ""),
    #    width = 800, height = 600)
    plot(
      country_data$year, 
      country_data$happiness, 
      type = "l", 
      main = paste("Trends for", country),
      xlab = "Year", 
      ylab = "Happiness Score", 
      col = "blue", 
      ylim = c(min(c(country_data$comb, country_data$neutral))
               , 8.5),
      lwd = 3
    )
    lines(country_data$year, country_data$in_one, col = "red", lwd = 2)
    legend(
      "topright", 
      legend = c("Happiness", "Combined Moods (all 3)"), 
      col = c("blue", "red"), 
      lty = 1, 
      lwd = 2,
      cex = 0.5
    )
    #dev.off()
  }
}

plot_country_trends(df)
```

```{r, echo=F}
write.csv(df, "df_used_in_R.csv", row.names = FALSE)
```

